---
---
---
title: "Supplemental Figures"
date: "Last knitted on `r format(Sys.Date(), '%d %b %Y')`"
author: "Ciara Asamoto, Sebastian Kopf"
output:
  html_document: 
    df_print: paged # omit to disable paged data table output
    css: stylesheet.css # omit if no need for custom stylesheet
    number_sections: no # change to no for unnumbered sections
    toc: yes # change to no to disable table of contents
    toc_float: true # change to false to keep toc at the top
    toc_depth: 3 # change to specify which headings to include in toc
    code_folding: hide # change to 'show' to show code by default
editor_options:
  chunk_output_type: console # change to inline to show output inline
---

```{r setup, echo = FALSE, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse) 
library(latex2exp)
library(readxl)
source(file.path("scripts", "plotting_functions.R"))
source(file.path("scripts", "table_functions.R"))
source(file.path("scripts", "growth_functions.R"))

# global knitting options for automatic saving of all plots as .png and .pdf
knitr::opts_chunk$set(
  dev = c("png", "pdf"), fig.keep = "all",
  dev.args = list(pdf = list(encoding = "WinAnsi", useDingbats = FALSE)),
  fig.path = file.path("figures", "figure_S")
)
```

# Load Data

> NOTE: please knit the `calculations.Rmd` notebook first to prepare the data used here.

```{r}
# load experiments info
experiments <- readxl::read_excel(file.path("data", "experiments.xlsx"))

# load strains info
strains <- readxl::read_excel(file.path("data", "strains.xlsx")) 

# load concentration data
conc_data_w_t0 <- read_rds(file.path("cache", "conc_data_w_t0.rds"))

# load isotope data
iso_data_w_t0 <- read_rds(file.path("cache", "iso_data_w_t0.rds"))

# load growth data
growth_data <- readxl::read_excel(file.path("data", "data.xlsx"), sheet = "growth_data")

# load growth rates data
growth_rates_data <- read_rds(file.path("cache", "growth_rates_data.rds"))

# fractionation factors
eps_data <- read_rds(file.path("cache", "eps_data.rds"))

# fractionationc coupling
coupling_data <- read_rds(file.path("cache", "coupling_data.rds"))
```

# Table S1: Expanded Data Summary

```{r}
# prepare data
all_data <- 
  experiments %>% 
  left_join(strains, by = c("strain_id")) %>% 
  right_join(eps_data, by = c("exp_id")) %>% 
  left_join(coupling_data, by = c("strain_id", "medium")) %>% 
  left_join(growth_rates_data, by = c("exp_id", "culture")) %>% 
  arrange(name, medium, exp_id, culture) %>% 
  transmute(
    Species = name,
    `Reductase genes` = nitrate_reductases,
    Medium = medium,
    Tracer = enriched_water,
    `18eps / 15eps` = sprintf("%.2f +/- %.2f", ON_ratio, ON_ratio_se),
    `15eps` = sprintf("%.1f +/- %.1f", eps15N.permil, eps15N_se.permil),
    `18eps` = sprintf("%.1f +/- %.1f", eps18O.permil, eps18O_se.permil),
    `Growth rate (1/hour)` = ifelse(!is.na(mu.1_hr), sprintf("%.2f +/- %.2f", mu.1_hr, mu_se.1_hr), NA),
  )

# save table
all_data %>% 
  mutate(
    same_species = c("DNE", head(Species, -1)) == Species,
    same_medium = same_species & c("DNE", head(Medium, -1)) == Medium,
    same_tracer = c("DNE", head(Tracer, -1)) == Tracer,
    Species = ifelse(same_species, NA_character_, Species),
    `Reductase genes` = ifelse(same_species, NA_character_, `Reductase genes`),
    Medium = case_when(
      same_medium & same_tracer ~ NA_character_,
      Tracer == "yes" ~ paste(Medium, "+", "18O water"),
      TRUE ~ Medium
    ),
    `18eps / 15eps` = ifelse(same_medium, NA_character_, `18eps / 15eps`)
  ) %>% 
  select(-Tracer, -starts_with("same_")) %>% 
  export_to_excel(file = file.path("tables", "table_S1.xlsx"))
  
all_data
```

# Prepare Plots

```{r "base_plot"}
# experiment and strain info combined
exp_strains <- 
  experiments %>% 
  left_join(
    select(strains, strain_id, name, nitrate_reductases), 
    by = c("strain_id")
  ) %>% 
  mutate(
    name = str_replace(name, "\U0394", "$\\\\Delta$"),
    panel = sprintf("%s (%s)", name, medium) %>% factor(),
    medium = paste("Medium:", medium),
  )

# generate plot template for most of the SI plots
base_plot <- 
  ggplot() +
  aes(color = as.factor(culture), shape = enriched_water) +
  facet_wrap(~panel, ncol = 3, scales = "free", labeller = latex_labeller, drop = FALSE) +
  theme_figure(text_size = 16, strip_text_size = 10) +
  scale_color_manual(values = cb_palette) +
  labs(color = "Culture", shape = latex2exp::TeX("$^{18}$O water added?"))
```

# Figure S1: Growth curves

```{r "2_OD_vs_time", fig.width=10, fig.height=10, warning=FALSE}
# prepare data frame for plotting
ps1_data <- 
  growth_data %>% 
  left_join(exp_strains, by = c("exp_id"))
ps1_fit <- 
  growth_rates_data %>% 
  left_join(exp_strains, by = c("exp_id")) %>% 
  generate_logistic_curve(
    time = hour, N = OD, N0 = OD0, 
    time_min = hour_min, time_max = hour_max,
    K = K, r = mu.1_hr
  )

# generate plot
ps1 <- 
  base_plot %+%
  ps1_data %+%
  aes(x = hour, y = OD) +
  # fit
  geom_line(
    data = ps1_fit, 
    mapping = aes(shape = NULL, group = paste(exp_id, culture))
  ) +
  # data
  geom_point(size = 3) +
  # labels
  labs(x = "Time (hours)", y = "OD")

ps1
```

# Figure S2: Nitrate & nitrite concentrations over time

```{r "2_nitrate_nitrite_vs_time", fig.width=10, fig.height=10, warning=FALSE}
# prepare data frame for plotting
ps2_data <- 
  left_join(conc_data_w_t0, iso_data_w_t0, by = c("exp_id", "culture", "timepoint")) %>%
  left_join(exp_strains, by = c("exp_id")) %>% 
  mutate(NO2.mM = ifelse(is.na(NO2.mM), 0, NO2.mM)) %>% 
  pivot_longer(c(NO3.mM, NO2.mM), names_to = "var", values_to = "val") %>% 
  filter(!is.na(hour), !is.na(val)) %>% 
  mutate(
    conc = factor(var) %>% fct_rev() %>% fct_recode(
      "$NO_2^-$ (mM)" = "NO2.mM",
      "$NO_3^-$ (mM)" = "NO3.mM"
    )
  )

# generate plot
ps2 <- 
  base_plot %+% 
  ps2_data %+%
  aes(x = hour, y = val, linetype = conc) +
  # data
  geom_line() +
  geom_point(size = 3) +
  # star IC purified data points
  geom_text(
    data = function(df) 
      filter(df, var == "NO2.mM", str_detect(treatment, fixed("IC"))) %>% 
      mutate(label = "*"),
    mapping = aes(label = label), size = 8, color = "black", 
    hjust = 0.5, vjust = 0.5, nudge_x = -1, nudge_y = -0.5
  ) +
  # scales and labels
  scale_linetype_manual(values = c(1, 2), labels = latex_labeller) + 
  labs(x = "Time (hours)", y = "Concentration", linetype = "Concentration")

ps2
```

# Figure S3: $\Delta\delta^{15}$N vs. $\ln f$

```{r "3_Dd15N_vs_lnf", fig.width=10, fig.height=10, warning=FALSE}
# prepare data frame for plotting (all cultures with more than 2 data points)
ps3_4_data <- 
  inner_join(conc_data_w_t0, iso_data_w_t0, by = c("exp_id", "culture", "timepoint")) %>%
  left_join(exp_strains, by = c("exp_id")) %>% 
  group_by(exp_id, culture) %>% filter(n() > 2L) %>% ungroup()

# generate plot
ps3 <- 
  base_plot %+% 
  ps3_4_data %+%
  aes(x = ln_NO3, y = D_d15N) +
  # linear regression fits
  geom_smooth(method = lm, formula = y ~ x, se = FALSE) +
  # data
  geom_point(size = 3) +
  # star IC purified data points
  geom_text(
    data = function(df) 
      filter(df, str_detect(treatment, fixed("IC"))) %>% 
      mutate(label = "*"),
    mapping = aes(label = label), size = 8, color = "black", 
    hjust = 0.5, vjust = 0.5, nudge_x = -0.1, nudge_y = -0.5
  ) +
  # scales and labels
  scale_x_reverse() +
  labs(
     x = latex2exp::TeX("ln(\\[NO_3^-\\] / \\[NO_3^-\\]_{initial})$"),
     y = latex2exp::TeX("$\\Delta \\delta^{15}N$ of  $NO_3^-$")
  )

ps3
```

# Figure S4: $\Delta\delta^{18}$O vs. $\ln f$

```{r "4_Dd18O_vs_lnf", fig.width=10, fig.height=10, warning=FALSE}
# generate plot
ps4 <- 
  base_plot %+% 
  ps3_4_data %+% # same data as in fig. S3
  aes(x = ln_NO3, y = D_d18O) +
  # linear regression fits
  geom_smooth(method = lm, formula = y ~ x, se = FALSE) +
  # data
  geom_point(size = 3) +
  # star IC purified data points
  geom_text(
    data = function(df) 
      filter(df, str_detect(treatment, fixed("IC"))) %>% 
      mutate(label = "*"),
    mapping = aes(label = label), size = 8, color = "black", 
    hjust = 0.5, vjust = 0.5, nudge_x = -0.1, nudge_y = -0.5
  ) +
  # scales and labels
  scale_x_reverse() +
  labs(
     x = latex2exp::TeX("ln(\\[NO_3^-\\] / \\[NO_3^-\\]_{initial})$"),
     y = latex2exp::TeX("$\\Delta \\delta^{18}O$ of  $NO_3^-$")
  )

ps4
```

# Figure S5: $\Delta\delta^{18}$O vs. $\Delta\delta^{15}$N

```{r "5_Dd18O_vs_Dd15N", fig.width=10, fig.height=10, warning=FALSE}
# prepare data frame for plotting
ps5_data <- 
  iso_data_w_t0 %>%
  left_join(exp_strains, by = c("exp_id"))

# generate plot
ps5 <- 
  base_plot %+% 
  ps5_data %+%
  aes(x = D_d15N, y = D_d18O) +
  # 0.5 and 1.0 reference lines
  geom_abline(
    data = tibble(
      icept = 0, slope = c(1, 0.5), 
      line = sprintf("%.1f", slope) %>% factor() %>% fct_inorder()
    ),
    mapping = aes(
      x = NULL, y = NULL, color = NULL, shape = NULL,
      intercept = icept, slope = slope, linetype = line
    )
  ) +
  # linear regression fits
  geom_smooth(
    mapping = aes(color = NULL, shape = NULL),
    method = lm, formula = y ~ x, se = FALSE, color = "red",
    show.legend = FALSE
  ) +
  # data
  geom_point(size = 3) +
  # star IC purified data points
  geom_text(
    data = function(df) 
      filter(df, str_detect(treatment, fixed("IC"))) %>% mutate(label = "*"),
    mapping = aes(label = label), size = 8, color = "black", 
    hjust = 0.5, vjust = 0.5, nudge_x = -1.5, nudge_y = -0.5
  ) +
  # scales and labels
  labs(
     x = latex2exp::TeX("$\\Delta \\delta^{15}N$ of  $NO_3^-$"),
     y = latex2exp::TeX("$\\Delta \\delta^{18}O$ of  $NO_3^-$"),
     linetype = latex2exp::TeX("$\\frac{^{18}\\epsilon}{^{15}\\epsilon}$ ratio")
  ) + 
  guides(color = guide_legend(override.aes = list(linetype = 0)))

ps5
```

# Figure S6: 18O water data

```{r "6_18O_experiments", fig.width = 9, fig.height = 5, warning=FALSE, message=FALSE}
# prepare data frame for plotting
ps6_data <- 
  iso_data_w_t0 %>%
  left_join(exp_strains, by = c("exp_id")) %>% 
  filter(enriched_water == "yes") %>% 
  mutate(
    panel = name %>% fct_drop() %>% { fct_relevel(., levels(.)[c(2,4,3,1)]) }
  )

# generate plot
ps6 <- 
  base_plot %+% 
  ps6_data %+%
  aes(
    x = D_d15N, y = D_d18O, 
    color = nitrate_reductases, shape = nitrate_reductases
  ) +
  # 0.5 and 1.0 reference lines
  geom_abline(
    data = tibble(
      icept = 0, slope = c(1, 0.5), 
      line = sprintf("%.1f", slope) %>% factor() %>% fct_inorder()
    ),
    mapping = aes(
      x = NULL, y = NULL, color = NULL, shape = NULL,
      intercept = icept, slope = slope, linetype = line
    )
  ) +
  # data
  geom_point(size = 4, alpha = 0.9) +
  # scales and labels
  scale_shape_manual(values = c(15, 17, 16)) +
  scale_colour_manual(
    values = c("#660099", "#0066CC", "#CC0000")
  ) +
  labs(
     x = latex2exp::TeX("$\\Delta \\delta^{15}N$ of  $NO_3^-$"),
     y = latex2exp::TeX("$\\Delta \\delta^{18}O$ of  $NO_3^-$"),
     linetype = latex2exp::TeX("$\\frac{^{18}\\epsilon}{^{15}\\epsilon}$ ratio"),
     color = "reductases", shape = "reductases"
  ) + 
  guides(color = guide_legend(override.aes = list(linetype = 0)))

ps6
```


